Dynamic strategies for yaw and induction control of wind farms based on large-eddy simulation and optimization

### 2018-01-07 reply_to_reviewer2.tex

``
In the current manuscript, we aim to fill some of the aforementioned gaps in wind-farm control research \textbf{using the dynamic control approach based on large-eddy simulation and optimization as introduced by Goit \& Meyers [11]. The main novelty of the current work is the inclusion of yaw control, in addition to the induction control considered in previous studies [11, 12, 14].} Based on a demonstration $\dots$
''

\hrulefill

\paragraph{2. Reviewer} \textit{Line 137: "...in an accurate model with full state...", references should be provided here with previous studies using this model to show its accuracy. Authors already cited them in Introduction but it would be worth to repeat some of them here.}

\paragraph{Response} Done. References to previous studies using the same model were added (\red{line XX, p YY}):

``
... accurate model with full state information (see, e.g. Refs. [11--14]), i.e. using large-eddy simulation ...
''

\red{WM: Heb hier gereferereerd naar onze vorige optimal controls met LES studies... Weet niet of hij dit exact bedoelt, of eerder naar in welke mate LES geschikt is voor wind-farm simulaties (dit vermelden we niet echt in de introductie, dus ben niet zeker...) Evt. te bespreken. }

@@

``
$\dots$, \revision{and the vertical location of the turbine hubs $z_h$ is placed at the vertical mid-plane of the domain, i.e. $z_h/L_z = 0.5$,} in order to minimize $\dots$
''

\hrulefill

\paragraph{7. Reviewer} \textit{Line 215-218: Figure should be provided with 3D domain, to clarify the vertical placement of the turbine.}

\paragraph{Response} Done, Figure 3 has been updated in the revised manuscript, and is shown above.

\begin{figure}
	\includegraphics[width=\textwidth]{figure3_revised}
\end{figure}

@@

\hrulefill

\paragraph{8. Reviewer} \textit{It might be good to provide brief comments on sensitivity of the solution to control time or explain the choice of current control time. If it was done previously, reference would be enough.}

\paragraph{Response} This is a valid comment. We have updated the discussion in the manuscript where time horizons $T$ and $T_A$ are selected as follows (\red{Section 3.1, line XX p YY}):

``
\revision{Within the context of the receding-horizon methodology introduced in Section 2, the optimization is performed over discrete time windows with optimization horizon $T$ and flow advancement time $T_A$. Ideally, one would take a very large optimization horizon $T$, and a small flow advancement time $T_A$ before recomputing the optimal control problem. However, in practice, $T$ is limited due to the inherent chaotic nature of turbulent flows that complicates control over long time horizons, and $T_A$ should not be taken too small to avoid excessive computational costs. In the current work, we use 6 consecutive windows, each with an optimization horizon $T=300$~s and a flow advancement time of $T_A = T/2= 150$~s, resulting in a total control time of 900~s. Note that, given a wind-farm length of $3 \times 6D = 1 800$ m and a free-stream velocity $U_\infty = 8$ m s$^{-1}$, the wind-farm flowthrough time is approximately 225 s. Hence, the total control time comprises about 4 flowthroughs, and the optimization horizon within each window covers about 1.33 flowthrough times. Because of the latter, each turbine row has the opportunity to take into account interaction with every downstream row, and the sensitivity of power gains to the optimization horizon $T$ is expected to be limited.} All time-averaged results ...
''

Furthermore, we quantified the sensitivity of attainable power gains to $T/T_A$ in previous work, and mention this explicitly in the revised manuscript (\red{Section 3.1, line XX p YY}):
`` 
... or all turbines. \revision{The sensitivity of attainable power gains to $T/T_A$, initial control guesses, and optimizer settings has been shown to be limited in a prior study [14].} Simulation and optimization parameters ...
''

@@

\hrulefill

\paragraph{9. Reviewer} \textit{It looks like no towers were included in current simulation. The authors should clearly state it in Introduction and provide comments on importance/non-importance of tower inclusion in Summary part. It could be provided after the authors provided comments on the effect of stratification.}

\paragraph{Response} Indeed, the turbines are modeled without adding a turbine tower. We mention this explicitly in the revised manuscript as (Section 2, \red{line XX, p YY}):

``
... non-rotating actuator disk model (see, e.g., Ref [38]). \revision{Turbine towers are not included.} For each ...
''

Prior studies in literature have shown that including towers and nacelles in wind-farm LES can improve the near-wake characteristics, although we expect that, at the grid resolution of the current paper, the improvement would be relatively small. However, we agree that in the future this would be a valuable improvement to the fidelity of the turbine model. Moreover, when considering loads as indicated in the future work section, assessment of tower loading is of high importance. As suggested by the reviewer, we include this in the Conclusions section as follows:

``
\revision{The current optimization studies have been performed using relatively simple non-rotating actuator disk models for the wind turbine rotors without the inclusion of towers or nacelles. Although it has been shown that, by using this approach, the fidelity of far wake characteristics is satisfactory, the credibility of the current power gains and control strategies in practice could be further increased through wind-tunnel testing and similar optimization studies with more detailed turbine models, e.g. by adding drag forces for the tower and nacelle or by representing the rotor with an actuator line model. These are left for future work.} Finally, note that the current study optimizes power at all costs, and did not account for any undesirable loading characteristics. These should also be considered upon thoroughly comparing induction and yaw control for practical purposes. \revision{In this latter case, a detailed turbine representation is of high importance.}
''

@@

\paragraph{Response} Thanks for pointing this out. Fixed.

\hrulefill

\paragraph{11. Reviewer} \textit{Line 335: The authors clearly stated that the optimal control is heterogeneous, however no comments on why it is heterogeneous were provided. It was stated that turbine can pick either wake meandering type of motion or wake redirection, however what will determine this decision?}

\paragraph{Response} As mentioned in the manuscript, the flow conditions of the different columns are equivalent in a \emph{statistical} sense. However, the gradient, and hence the updates to optimized control iterates, are based on \emph{instantaneous} flow conditions. Small instantaneous differences in flow conditions between different columns can hence tip the balance between different flow control mechanisms suchs as wake meandering, wake redirection through clockwise yawing, or wake redirection through counterclockwise yawing. Which mechanism is chosen depends on the initial steps in the optimization process when the optimizer is settling into a local optimum of the cost functional landscape. These steps are based on the gradient, which is found by solving the turbulent adjoint equations. Therefore, it is not possible to determine \emph{a priori} which method will be chosen by the optimizer. The fact that, after some time, all turbines seem to abandon the meandering mechanism in favor of the redirection mechanism seems to indicate that a global optimization approach would probably always yield the wake redirection mechanism.

We included the reason for different mechanisms in the optimized controls as follows (Section 4.2.2, \red{line XX, p YY}):
``
... their behavior in an optimal control setting is heterogeneous. \revision{The reason for this is that the cost functional gradient used in the optimization process is based on the instantaneous flow conditions through the adjoint equations. These conditions are not the same for different columns, e.g. the turbulence in the wakes behind the third row observed in Figure 5 is different for every column, which can cause the optimizer to converge to a different dominant control mechanism in different columns.} For example, ...
''

@@

\hrulefill

\paragraph{12. Reviewer} \textit{Figure 5: The blue color for C4 and green color for C2 very hard to see, especially on I3Y case.}

\paragraph{Response} Thank your for this justified comment. We have plotted all turbine locations in black in the revised version of Figure 5 (and the corresponding Figure 11 for turbulent inflow).

@@

\hrulefill

\paragraph{13. Reviewer} \textit{Figure 6 shows that for uniform inflow in a reference case the turbine in a second row will be producing almost 90\% less power then upstream. As a reader I would be interested to see previous studies on this. Adding corresponding references would be enough.}

\paragraph{Response} Indeed, power losses in uniform inflow conditions can be very large due to artificially stabilized wakes. In the submitted manuscript, we already mentioned that such conditions are somewhat artificial. We have expanded this discussion with some references in the revised manuscript. Furthermore, we added some references that also quantify severe losses under uniform inflow. A direct comparison between our results and prior work is not possible due to the dependence on numerical solver details, as indicated in the revised manuscript. The manuscript has been updated as follows (Section 3.1 \red{line XX, p YY}):

``
... These flow conditions can be regarded as somewhat artificial since, in the atmospheric boundary layer, background turbulence typically causes instant wake transition behind the turbine rotor. \revision{Furthermore, it is documented in literature that numerical details such as grid resolution, subgrid-scale modeling, and turbine models can have a significant impact on simulation results , whereas in turbulent inflow this is much less the case  [49-51]. In practice, this tends to lead to artificially stabilized wakes with very low farm efficiencies (see, e.g., Refs [52,53]).} Nevertheless, ...
''

Also, when discussing the power extraction of the refernece case in uniform inflow, we refer to this discussion (Section 4.2.1 \red{line XX, p YY}):
``
 As expected \revision{from the discussion in Section 3.1}, the efficiency of the uncontrolled reference farm R is very low at around $40\%$. ...
''

@@

\hrulefill

\paragraph{14. Reviewer} \textit{Figure 7: Add tick labels to X-axis, i.e. to Time. (Same for Fig. 13)}

\paragraph{Response} We thank the reviewer for pointing out the missing labels, this has been corrected in the revised versions of Figures 7 and 13. Note that, in response to comments from another reviewer, we have reduced the amount of information in these figures somewhat for better clarity.

### 2018-01-07 reply_to_reviewer3.tex

\paragraph{Reviewer} \textit{This paper presents an interesting study on yaw and induction control of wind farms. The analysis is reliable, and the paper is well written.}

\section*{Comments}
\hrulefill

\paragraph{1. Reviewer} \textit{Page 11, line 325, “additional control cases Ystat and Ymndr…” Although explanations of Ystat and Ymndr can be found later, e.g. on page 20, line 521. It is suggested that these control cases are explained upon their first appearance. In the caption of Figure 6, the case names should be noted as well.}

\paragraph{Response} We thank the reviewer for this comment. Agreed, a definition of these cases upon their first appearance is required. We briefly explained the definitions of these cases as follows (\red{line XX, p YY}):

`` 
The figure also includes results for two additional control cases that are based on simplified controls derived from the yaw characteristics of the optimal control case Y. \revision{Firstly, Ystat denotes a static yaw control case based upon the time-averaged yaw angles of case Y. Secondly, Ymndr is a dynamic yaw control case in which the first-row turbines performs small yaw rotations in an alternating directions.} These simplified control cases are further detailed and discussed below.
''

A more detailed explanation is given during the analysis of the yaw signals, as the definition of these simplified cases is based upon this analysis.

The caption of Figure 6 is updated with the casenames as follows:

`` 
Figure 6: \revision{Wind-farm power extraction for reference case R, optimal control cases (Y, Y5, I2, I3, I2Y, and I3Y), and derived control cases (Ystat and Ymndr, see further description below)}. \emph{a) } Row-averaged power, normalized by first-row power of reference case R. \emph{b) } Wind-farm efficiency $\eta_{\text{farm}}$ ... 
''

@@

\hrulefill

\paragraph{2. Reviewer} \textit{Some figures are difficult to inspect closely. For example, Figures 7, 11, and 13 contain color lines, and the sizes are too small. The authors can present the results in a better manner.}

\paragraph{Response}

\begin{itemize}
	\item Figure 11 (and Figure 5): due to the colored turbines, some turbines were hard to see against the colored background image. The turbines have been plotted in black in the revised figures. 
	\item Figure 7 and 13: we agree that these figures were hard to read, but they contained a lot of information that was further used in the analysis. As a compromise, we plot the results only for column C2 and C4, and omit columns C1 and C3. This allows us to still distinguish between the redirecting and meandering regimes, and reduces the amount of clutter in the figure. Because of this, also some  minor adaptations in the manuscript text had to be made when describing the yaw time evolution. (see revised manuscript) Furthermore, Figures 5 and 11 have been plotted for $t=300$~s instead of the initial value of 450 s: since C2 in case Y has switched from meandering to redirection at 450~s, this could be confusing to the reader. 
\end{itemize}

@@

\hrulefill

\paragraph{3. Reviewer} \textit{In Figures 9 and 15, the unit of PSD should be added.}

\paragraph{Response} The power spectral density has units of ${\rm rad}^2 / {\rm Hz}$. However, in order to be consistent with the non-dimensional x-axis (i.e. frequency in the form of a Strouhal number), we also non-dimensionalized the PSDs appropriately as ${\rm PSD}_\theta \ U_\infty / D \sigma_{\theta}^2$. \red{Also note that, in response to the comments from another reviewer, we have not detrended the data in the revised figures, affecting the DC component of the spectra in this figure.}

@@

\hrulefill

\paragraph{4. Reviewer} \textit{Table 2 gives a summary of the effect different control strategies. Why does the meandering yaw case have reduced power gains in the turbulent condition? How do the authors judge the uncertainties in their analysis? For example, by extending the optimization time window or by refining the cell size, will the conclusion be different?}

\paragraph{Response}

\begin{itemize}
	\item \textit{Why does the meandering yaw case haw reduced power gains in the turbulent conditions?} \\ We believe this is answered properly in the text (\red{l XX, p YY}):\\ `` However, in turbulent inflow conditions the same strategy did not lead to an increase in power extraction. This can be explained by the fact that meandering is already triggered naturally in turbulent boundary layers. Instead of actively reinforcing downstream wake meandering, the faint traces of spectral peaks at $St = 0.2$ observed in Figure~15 might therefore simply be the response of yawing turbines to variability in the incident flow angle induced by natural upstream wake meandering. Another hypothesis is that reinforcing natural meandering requires more complex control of the yaw rate than the currently considered bang-bang control, and that the phase of yaw angle variations might have to be matched to local flow conditions. Further research is required to confirm or negate these hypotheses.''
	\item \textit{How do the authors judge the uncertainties in their analysis?} \\ Indeed, the optimization setup requires some user-defined parameters. In an earlier study [14], we have performed a parameter study for some of them, leading to the currently used values for e.g. the initial control guess, amount of retained Hessain pairs in the BFGS iterations, and $T/T_A$, and were able to confirm that the choices made for these parameters do not significantly affect the outcome of the simulations. We added this as a remark in the manuscript as follows (\red{line XX, p YY}):\\ ``... for all turbines. \revision{The sensitivity of attainable power gains to $T/T_A$, initial control guesses, and optimizer settings has been shown to be limited in a prior study[14].} Simulation ...''\\ Regarding the specific examples given by the reviewer:
	\begin{enumerate}
		\item \textit{Extending the optimization time window}: A longer optimization horizon $T$ leads to more opportunity for interaction between the turbines and hence potentially larger power gains. However, given the limited wind-farm size in the current study, we are able to use an optimization horizon (300 s) well over the wind-farm flowthrough time ($\approx 225$ s). This means that interactions between each of the turbine rows can be taken into account during the optimization for a significant portion of the window. Therefore, the sensitivity of the conclusions to extending the optimization window is expected to be small. This has been mentioned in the manuscripts as follows (\red{line XX, p YY}): \\``...resulting in a total control time of 900~s. \revision{Note that, given a wind-farm length of $3 \times 6D = 1 800$ m and a free-stream velocity $U_\infty = 8$ m s$^{-1}$, the wind-farm flowthrough time is approximately 225 s. Hence, the total control time comprises about 4 flowthroughs, and the optimization horizon within each window covers about 1.33 flowthrough times. Because of the latter, each turbine row has the opportunity to take into account interaction with every downstream row, and the sensitivity of power gains to the optimization horizon $T$ is expected to be limited.} All time-averaged results ...''
		\item \textit{Refining the cell size}: In order to limit computational costs, we use a relatively limited resolution to run the optimal control cases. In earlier work [1], a numerical experiment was performed to assess the grid sensitivity of a rotating ADM using a solver with similar pseudo-spectral / finite-difference numerics as our solver. It was found that, for resolutions close to the one used in the current study (our resolution is slightly higher than case ADM-R3 in the mentioned study), simulation results were relatively insensitive to grid resolution, except for some small discrepancies in the near-wake top-tip turbulence intensity. Therefore, we are confident about the grid sensitivity of the conclusions in the current study. \red{WM - Weet niet of we dit expliciet in de tekst moeten verwerken? evt. te bespreken...}
	\end{enumerate}
\end{itemize}

[1] Y.-T. Wu, F. Port\'e-Agel. ``Large-eddy simulation of wind-turbine wakes: evaluation of turbine parametrisations'' Boundary-Layer Meteorology 128 (2011): 345--366
